### 2015-03-31 Kokkos_PG_Compilation.tex
Convert to SAND format.

\chapter{Compiling}\label{C:build}

@@

\section{General Information}\label{S:build:gen}

@@

\section{Using Kokkos' Makefile system}\label{S:build:make}

@@

\section{Using Trilinos' CMake build system}\label{S:build:Trilinos}

@@

\section{Building for CUDA}\label{S:build:CUDA}

@@

\section{Building for Intel Xeon Phi}\label{S:build:Phi}

>>>>

\section{Compiling}\label{C:build}

@@

\subsection{General Information}\label{S:build:gen}

@@

\subsection{Using Kokkos' Makefile system}\label{S:build:make}

@@

\subsection{Using Trilinos' CMake build system}\label{S:build:Trilinos}

@@

\subsection{Building for CUDA}\label{S:build:CUDA}

@@

\subsection{Building for Intel Xeon Phi}\label{S:build:Phi}

### 2015-03-31 Kokkos_PG_HierarchicalParallelism.tex
Convert to SAND format.

\chapter{Hierarchical Parallelism}\label{C:Hierarchical}

@@

\section{Motivation}\label{S:Hierarchical:Motivation}

@@

\section{Thread teams}\label{S:Hierarchical:Teams}

@@

\subsection{Creating a Policy instance}\label{SS:Hierarchical:Teams:Policy}

@@

\subsection{Basic kernels}\label{SS:Hierarchical:Teams:Kernels}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Team scratch pad memory}\label{S:Hierarchical:Scratch}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nested parallelism}\label{S:Hierarchical:Nested}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Team and vector loops}\label{SS:Hierarchical:Nested:Loops}

@@

\subsection{Restrict execution with \lstinline|single|}

>>>>

\section{Hierarchical Parallelism}\label{C:Hierarchical}

@@

\subsection{Motivation}\label{S:Hierarchical:Motivation}

@@

\subsection{Thread teams}\label{S:Hierarchical:Teams}

@@

\subsubsection{Creating a Policy instance}\label{SS:Hierarchical:Teams:Policy}

@@

\subsubsection{Basic kernels}\label{SS:Hierarchical:Teams:Kernels}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Team scratch pad memory}\label{S:Hierarchical:Scratch}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nested parallelism}\label{S:Hierarchical:Nested}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Team and vector loops}\label{SS:Hierarchical:Nested:Loops}

@@

\subsubsection{Restrict execution with \lstinline|single|}

### 2015-03-31 Kokkos_PG_Initialization.tex
Convert to SAND format.

\chapter{Initialization}\label{C:init}

@@

\section{Initialization by command-line arguments}\label{S:init:cmdline}

@@

\section{Initialization by struct}\label{S:init:struct}

@@

If you set \lstinline|num_threads| or \lstinline|num_numa| to zero or less, 
Kokkos will \emph{not} initialize the default host execution space.
Note that the default value of each parameter is -1.
Thus, you must set both \lstinline|num_threads| and \lstinline|num_numa|
in order for Kokkos to initialize the default host execution space.

@@

\section{Initializing non-default execution spaces}\label{S:init:specific}

@@

\section{Finalization}\label{S:init:finalize}

>>>>

\section{Initialization}\label{C:init}

@@

\subsection{Initialization by command-line arguments}\label{S:init:cmdline}

@@

\subsection{Initialization by struct}\label{S:init:struct}

@@

If you set \lstinline|num_threads| or \lstinline|num_numa| to zero or less, 
Kokkos will try to determine default values if possible or otherwise set them to 1.
In particular Kokkos can use the hwloc library to determine default settings, using the assumption that the process binding mask is unique, i.e. that this process does not share any cores with another process.
Note that the default value of each parameter is -1.

@@

\subsection{Initializing non-default execution spaces}\label{S:init:specific}

@@

\subsection{Finalization}\label{S:init:finalize}

### 2015-03-31 Kokkos_PG_Introduction.tex
Convert to SAND format.

\chapter{Introduction}\label{C:intro}

\noindent
The field of High Performance Computing (HPC) is on the verge of entering a new era. The need
for a fundamental change comes from many angles including the growing acceptance that rates of
pure computation (often called FLOP/s) are a poor single optimization goal for scientific
workloads as well as practical challenges in the form of producing energy efficient and cost
efficient processors. Since the convergence on the Message-Passing Interface (MPI) standard
in the mid 1990s, application developers have enjoyed a fairly static view of the machine --
that of a distributed collection of homogeneous nodes executing in collaboration. However, 
after almost two decades of domaince, the sole use of MPI to derive parallelism is acting as
a potential limiter to greater future performance. While we expect MPI to continue to function
as the basic mechanism for communication between compute nodes, additional parallelism is
likely to be required on the computing node itself if high performance and efficiency are to be
realized.

In reviewing potential options for the computing nodes of the future the reader might fall
upon three broad categories of computing device: (1) the multi-core processor with powerful
serial performance, optimized to reduce operation latency; (2) many-cores processors with
low to medium powered compute cores that are designed to extent the multi-core concept toward
throughput based computation and, finally, (3), the general-purpose graphics processor unit
(GP-GPU, or often, GPU) which is a much more numerous collection of low powered cores designed
to tolerate long latencies but provide performance through a much higher degree of 
parallelism. Any combination of these options might also be combined in the future.

The diversity of the options available for processor selection raises as interested question
as to how these should be programmed. Partly due to their heritage but also due to their
optimized designs, each of these hardware types offers a different programming solution and
a different set of guidelines to write applications to for highest performance. Options available
today include a number of shared memory approaches such as OpenMP, Cilk+, Thread Building Blocks
(TBB) as well as Linux p-threads. To target both contemporary multi/many-core processors and
GPUs technologies such as OpenMP, OpenACC and OpenCL might be used. Finally, for GPUs a 
programming model such as CUDA may be selected. This poses a problem to the application
developer of today which is reminiscent of the challenges before MPI became the default
communication library -- which model should be selected to provide portability across hardware
solutions as well as provide reasonable to high performance across each class of processor and
protect algorithm investment into the future. So far none of these models have been able to 
provide practical solutions to these questions.

The Kokkos programming model described in this programming guide seeks to address these
concerns by providing an abstraction of both computation and application data allocation and
layout. These abstraction layers are designed specifically to isolate software developers for
the fluctuation in hardware details yet provide portability and high levels of performance
across architectures.

>>>>

\section{Introduction}
\label{C:intro}

\noindent
The field of High Performance Computing (HPC) is on the verge of entering a new era. The need
for a fundamental change comes from many angles including the growing acceptance that rates of
pure computation (often called FLOP/s) are a poor single optimization goal for scientific
workloads, as well as practical challenges in the form of producing energy efficient and cost
efficient processors. Since the convergence on the Message-Passing Interface (MPI) standard
in the mid 1990s, application developers have enjoyed a seemingly static view of the 
underlying machine --
that of a distributed collection of homogeneous nodes executing in collaboration. However, 
after almost two decades of domaince, the sole use of MPI to derive parallelism is acting as
a potential limiter to greater future performance. While we expect MPI to continue to function
as the basic mechanism for communication between compute nodes for the immediate future, 
additional parallelism is
likely to be required on the computing node itself if high performance and efficiency goal 
are to be realized.

In reviewing potential options for the computing nodes of the future the reader might fall
upon three broad categories of computing device: (1) the multi-core processor with powerful
serial performance, optimized to reduce operation latency; (2) many-core processors with
low to medium powered compute cores that are designed to extend the multi-core concept toward
throughput based computation and, finally, (3), the general-purpose graphics processor unit
(GP-GPU, or often, GPU) which is a much more numerous collection of low powered cores designed
to tolerate long latencies but provide performance through a much higher degree of 
parallelism and computational throughput. Any combination of these options might also be 
combined in the future.

The diversity of the options available for processor selection raises as interested question
as to how these should be programmed. In part due to their heritage, but also due to their
optimized designs, each of these hardware types offers a different programming solution and
a different set of guidelines to write applications to for highest performance. Options available
today include a number of shared memory approaches such as OpenMP, Cilk+, Thread Building Blocks
(TBB) as well as Linux p-threads. To target both contemporary multi/many-core processors and
GPUs technologies such as OpenMP, OpenACC and OpenCL might be used. Finally, for highest
performance on GPUs a 
programming model such as CUDA may be selected. Such a variety of options poses a problem 
to the application
developer of today which is reminiscent of the challenges before MPI became the default
communication library -- which model should be selected to provide portability across hardware
solutions, as well as provide high performance across each class of processor and
protect algorithm investment into the future. None of the models listed above have been able to 
provide practical solutions to these questions.

The Kokkos programming model described in this programming guide seeks to address these
concerns by providing an abstraction of both computation and application data allocation and
layout. These abstraction layers are designed specifically to isolate software developers from
fluctuation and diversity in hardware details yet provide portability and high levels of performance
across many architectures.

### 2015-03-31 Kokkos_PG_LegacyCodeIntegration.tex
Convert to SAND format.

It is important to recognize that this does not lift certain restrictions. 
For example it is not valid to allocate a view inside of an OpenMP 
parallel region, the same way as it is not valid to allocate a View 
inside a \lstinline|parallel_for| kernel. Indeed there are things which 
are slightly more cumbersome when mixing the models. Assigning 
one view to another inside a Cuda kernel or an OpenMP parallel 
region is only possible if the destination view is unmanaged. 
During dispatch of kernels with \lstinline|parallel_for|  all Views referenced 
in the functor or lambda are automatically switched into unmanaged 
mode. This would not happen when simply entering an OpenMP 
parallel region.

@@

\subsection{Raw allocations through Kokkos}

@@

Inside of parallel sections \lstinline|Kokkos::vector| switches to view semantics. That means
in particular that assignments are shallow copies. Certain functions will also throw runtime 
errors when called inside a parallel kernel. This includes resize and push.

\section{Calling non-Kokkos libraries}

>>>>

It is important to recognize that this does not lift certain restrictions. 
For example it is not valid to allocate a view inside of an OpenMP 
parallel region, the same way as it is not valid to allocate a View 
inside a \lstinline|parallel_for| kernel. Indeed there are things which 
are slightly more cumbersome when mixing the models. Assigning 
one view to another inside a Cuda kernel or an OpenMP parallel 
region is only possible if the destination view is unmanaged. 
During dispatch of kernels with parallel for  all Views referenced 
in the functor or lambda are automatically switched into unmanaged 
mode. This would not happen when simply entering an OpenMP 
parallel region.

@@

\subsubsection{Kokkos manages memory}

\paragraph{Raw allocations}

@@

\subsection{Calling non-Kokkos libraries}

### 2015-03-31 Kokkos_PG_MachineModel.tex
Convert to SAND format.

\chapter{Machine Model}
\label{C:Model}

@@

\section{Motivations}\label{S:Model:Motivations}

@@

\subsection{Kokkos Abstract Machine Model}

@@

\subsection{Kokkos Memory Spaces}

@@

\subsubsection{Atomic accesses to Memory in Kokkos}

@@

\subsubsection{Memory Consistency in Kokkos}

@@

\section{Thread teams}\label{S:Model:Teams}

@@

\section{Program execution}\label{S:Model:Exec}

@@

\subsection{Reproducible reductions and scans}\label{SS:Model:Exec:Repro}

@@

\subsection{Asynchronous parallel dispatch}\label{SS:Model:Exec:Async}

@@

\subsection{Thread safety?}\label{SS:Model:Exec:ThreadSafety}

>>>>

\section{Machine Model}
\label{C:Model}

@@

\subsection{Motivations}\label{S:Model:Motivations}

@@

\subsubsection{Kokkos Abstract Machine Model}

@@

\subsubsection{Kokkos Memory Spaces}

@@

\paragraph{Atomic accesses to Memory in Kokkos}

@@

\paragraph{Memory Consistency in Kokkos}

@@

\subsection{Thread teams}\label{S:Model:Teams}

@@

\subsection{Program execution}\label{S:Model:Exec}

@@

\paragraph{Reproducible reductions and scans}\label{SS:Model:Exec:Repro}

@@

\paragraph{Asynchronous parallel dispatch}\label{SS:Model:Exec:Async}

@@

\subsubsection{Thread safety?}\label{SS:Model:Exec:ThreadSafety}

### 2015-03-31 Kokkos_PG_Subviews.tex
Convert to SAND format.

\chapter{Subviews}\label{C:Subviews}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A subview is a slice of a View}\label{S:Subviews:Slice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to take a subview}\label{S:Subviews:How}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{C++11 type deduction}\label{SS:Subviews:How:auto}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dimension of a subview}\label{SS:Subviews:How:dim}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Degenerate Views}\label{SS:Subviews:How:degen}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\section{Subviews}\label{C:Subviews}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A subview is a slice of a View}\label{S:Subviews:Slice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{How to take a subview}\label{S:Subviews:How}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{C++11 type deduction}\label{SS:Subviews:How:auto}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Dimension of a subview}\label{SS:Subviews:How:dim}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Degenerate Views}\label{SS:Subviews:How:degen}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-03-31 Kokkos_PG_Views.tex
Convert to SAND format.

\chapter{View: Multidimensional array}\label{C:View}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Why Kokkos needs multidimensional arrays}\label{S:View:Why}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Creating and using a View}\label{S:View:CreateUse}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Constructing a View}

@@

\subsection{What types of data may a View contain?}

@@

\subsection{Const Views}

@@

\subsection{Accessing entries (indexing)}

@@

\subsection{Reference counting}

@@

\subsection{Resizing}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Layout}\label{S:View:Layout}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Strides and dimensions}\label{SS:View:Layout:Strides}

@@

\subsection{Other layouts}\label{SS:View::Layout::Other}

@@

\subsection{Default layout depends on execution space}\label{SS:View:Layout:Default}

@@

\subsection{Explicitly specifying layout}\label{SS:View:Layout:Explicit}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Managing Data Placement}\label{S:View:Placement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Memory spaces}

@@

\subsection{Initialization}

@@

\subsection{Deep copy and HostMirror}

@@

\subsection{How do I get the raw pointer?}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Access traits}\label{S:View:AccessTraits}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Atomic access}

@@

\subsection{Random Access}

@@

\subsection{Standard idiom for specifying access traits}

@@

\subsection{Unmanaged Views}

>>>>

\section{View: Multidimensional array}\label{C:View}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Why Kokkos needs multidimensional arrays}\label{S:View:Why}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Creating and using a View}\label{S:View:CreateUse}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Constructing a View}

@@

\subsubsection{What types of data may a View contain?}

@@

\subsubsection{Const Views}

@@

\subsubsection{Accessing entries (indexing)}

@@

\subsubsection{Reference counting}

@@

\subsubsection{Resizing}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Layout}\label{S:View:Layout}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Strides and dimensions}\label{SS:View:Layout:Strides}

@@

\subsubsection{Other layouts}\label{SS:View::Layout::Other}

@@

\subsubsection{Default layout depends on execution space}\label{SS:View:Layout:Default}

@@

\subsubsection{Explicitly specifying layout}\label{SS:View:Layout:Explicit}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Managing Data Placement}\label{S:View:Placement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Memory spaces}

@@

\subsubsection{Initialization}

@@

\subsubsection{Deep copy and HostMirror}

@@

\subsubsection{How do I get the raw pointer?}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Access traits}\label{S:View:AccessTraits}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Atomic access}

@@

\subsubsection{Random Access}

@@

\subsubsection{Standard idiom for specifying access traits}

@@

\subsubsection{Unmanaged Views}
